---
title: "eiGetValues_Visualizer"
author: "Magie Williams"
date: "`r Sys.Date()`"
output: html_document
---
Priror to the PCA visualization varied bash scripts were used on an HPCC cluster to process data.

The PCA is from the PLINK PCA function after snps were genotyped with GATK and filtered with BCF Tools. 

**PCA Visualization**
```{r Defining Labels and Colors, echo=FALSE}
#Defining Labels and Colors
pop_colors <- c(
  "Port Sheldon" = "#FE6100",
  "Rosy Mound" = "#FFB000",
  "Seven Bridges" = "#0057B8",
  "Springvale" = "#DC267F",
  "Tatum" = "#785EF0"
)

pop_labels <- c(
  "Port Sheldon" = "Port Sheldon",
  "Rosy Mound" = "Rosy Mound",
  "Seven Bridges" = "Seven Bridges",
  "Springvale" = "Springvale",
  "Tatum" = "Tatum"
)

```

```{r Package Library}
#install if needed
library(readr)
library(dplyr)
library(ggplot2)
```

```{r Fixing Data, message=FALSE, warning=FALSE}
#I manually downloaded the files from the HPCC
eVector <- read_table("BWA_Out_quick.g.vcf.gz_pca.eigenvec", col_names = FALSE)
# Name columns: FID IID PC1..PC10
colnames(eVector) <- c("FID","IID", paste0("PC", 1: (ncol(eVector)-2)))
# Read eigenvalues and compute % variance
eValues <- read_table("BWA_Out_quick.g.vcf.gz_pca.eigenval", col_names = FALSE)[[1]]
var_pct <- 100 * eValues / sum(eValues)
pc1_lab <- sprintf("PC1 (%.2f%%)", var_pct[1])
pc2_lab <- sprintf("PC2 (%.2f%%)", var_pct[2])
#population IDs Clean
Pops <- read_csv("UpdatedSampleIDs.csv")
eVector <- eVector %>% left_join(Pops, by = "FID")
```


```{r First Plot, warning=FALSE}
gg <- ggplot(eVector, aes(x = PC1, y = PC2)) +
  geom_point(aes(color = Pop), size = 2.2) +
  theme_bw(base_size = 12) +
  labs(
    title = "PCA: PC1 vs PC2",
    x = pc1_lab,
    y = pc2_lab,
    color = "Population"
  ) +
  theme(legend.position = "right") +
  scale_color_manual(
    values = pop_colors,
    labels = pop_labels,
    name = "Population"
  )
print(gg)
#ggsave("TM_PC11_PC12.png", gg, width = 6.5, height = 5.5, dpi = 300)

# Add labels (optional)
# gg + ggrepel::geom_text_repel(aes(label = IID), size = 3)
```
While we may have the first plot, I like to double check my data distributions to make sure everything is looking okay, or if there is something I should flag. 

```{r Scree Plot, warning=FALSE}
#code generated with ai assistance for optimization
#need to run above chunk first
k <- min(13, length(var_pct))

scree <- data.frame(PC = factor(1:k), VarPct = var_pct[1:k])

s_gg <- ggplot(scree, aes(x = PC, y = VarPct)) +
  geom_col(fill = "#4C72B0") +
  geom_text(aes(label = sprintf("%.1f%%", VarPct)), vjust = -0.2, size = 3) +
  theme_bw(base_size = 12) +
  labs(title = "PCA Scree Plot", x = "Principal component", y = "Variance explained (%)")
print(s_gg)
#ggsave("TM_scree.png", s_gg, width = 6.5, height = 4, dpi = 300)

```
If everything looks good with the disctribution I shouldn't need much more plotting than the scree so I move onto prelime stats

```{r Stats Check, message=FALSE, echo=TRUE}
#anovas are the most appropriate so that we can do a Tukey
#Making Anova models first
PC1_anova <- aov(PC1 ~ Pop, data = eVector)
PC1_check <- summary(PC1_anova)
print(PC1_check)
PC2_anova <- aov(PC2 ~ Pop, data = eVector)
PC2_check <- summary(PC2_anova)
print(PC2_check)
#Run Tukey Test
tukey_PC1 <- TukeyHSD(PC1_anova)
print(tukey_PC1)
tukey_PC2 <- TukeyHSD(PC2_anova)
print(tukey_PC2)
#View Tukey
plot(tukey_PC1)
plot(tukey_PC2)
```


