---
title: "CleanedPixyResults"
author: "Magie Williams"
date: "`r Sys.Date()`"
output: pdf_document
---

I just wanted to clean up the results from pixy, since my other Rmarkdown is kind of just copy and paste different parts of code from me and copilot. I did use copilot to help with some of the cleaning and condensing. 

I did initially mess this up so I am updating as I go but this is the documentation notes from pixy:
"If you want to combine information across windows after the fact (e.g. by averaging), do not simply average the per-window summary statistics. Instead, sum the raw counts or components and recompute the statistic."

Before doing anything, I want to make sure everything is consistent between my Rmarkdowns so I am going to define labels and colors now then will get into the workflow. This is mostly for ease of coding late as well. 

Colorblind Friendly Colors from AudioEye: Black (#000000)

Orange (#E69F00) #FE6100
Sky Blue (#56B4E9)
Bluish Green (#009E73)
Yellow (#F0E442)
Blue (#0072B2)
Vermillion (#D55E00)
Reddish Purple (#CC79A7)

```{r}
pop_colors <- c(
  "Port Sheldon" = "#FE6100",
  "Rosy Mound" = "#FFB000",
  "Seven Bridges" = "#0057B8",
  "Springvale" = "#DC267F",
  "Tatum" = "#785EF0"
)

pop_labels <- c(
  "Port Sheldon" = "Port Sheldon",
  "Rosy Mound" = "Rosy Mound",
  "Seven Bridges" = "Seven Bridges",
  "Springvale" = "Springvale",
  "Tatum" = "Tatum"
)

sub_labels <- c("N" = "B. laevigata",
                "S" = "B. stricta")

sub_colors <- c( "N" = "")

```


```{r Package_Data, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(ggplot2)
#Read in pi file
pi <- read.delim("all_pops_pi.txt")
#adjust it to be usable
pi <- pi %>%
  mutate(
    chr_group = gsub("S$", "", chromosome),
    subgenome = ifelse(grepl("S$", chromosome), "S", "N")
  )

ggplot(pi, aes(x =avg_pi)) + 
  geom_histogram() 
#not weighted
ggplot(pi, aes(x = pop, y =avg_pi, fill = subgenome)) + 
  geom_boxplot() +
  ylab(expression(pi[Average] ("Not Weighted"))) +
  xlab("Population")
#whole genome investigation
#avg_pi -> this is not weighted
#amount of instancs where avg pi is zerp
sum(pi$avg_pi == 0, na.rm = TRUE)
#those grouped by population
ZeroPop <- pi %>%
  group_by(pop) %>%
  summarize(n_zero = sum(avg_pi == 0, na.rm = TRUE))
#grouped by chromosome
ZeroChr <- pi %>%
  group_by(chromosome) %>%
  summarize(n_zero = sum(avg_pi == 0, na.rm = TRUE))

mean(pi$avg_pi == 0, na.rm = TRUE)

sum(pi$avg_pi == 0, na.rm = TRUE) /
sum(!is.na(pi$avg_pi))

summary(pi$no_sites)
#positions with no sites
NoSites <- pi %>%
  filter(avg_pi == 0) %>%
  summarize(
    mean_sites = mean(no_sites, na.rm = TRUE),
    median_sites = median(no_sites, na.rm = TRUE)
  )
#self explanatory
CountCompare <- pi %>%
  filter(avg_pi == 0) %>%
  summarize(
    mean_comp = mean(count_comparisons, na.rm = TRUE),
    median_comp = median(count_comparisons, na.rm = TRUE)
  )
#also self explanatory
ZeroVsNonZero <- pi %>%
  summarize(
    total = n(),
    zero_pi = sum(avg_pi == 0, na.rm = TRUE),
    nonzero_pi = sum(avg_pi > 0, na.rm = TRUE),
    no_data = sum(is.na(avg_pi))
  )

#proportion_zero
PropZeroPop <- pi %>%
  group_by(pop) %>%
  summarize(
    prop_zero = mean(avg_pi == 0, na.rm = TRUE)
  )
#not weighted
NoZero <- pi %>%
  filter(avg_pi != "0")

range(NoZero$avg_pi)

ggplot(NoZero, aes(x =avg_pi)) + 
  geom_histogram() +
  labs(title = "Zeroes Removed")

ggplot(NoZero, aes(x = pop, y =avg_pi, fill = subgenome)) + 
  geom_boxplot() +
  ylab(expression(pi[Average] ("Not Weighted"))) +
  xlab("Population") +
  labs(title = "Zeroes Removed")
```
Okay so breaking down all the Pi Info:
Here is the number of observations where the avg_pi by window is zero

```{r, echo=FALSE} 
ZeroPop 
```

Now by chromosome instead

```{r, echo=FALSE}
ZeroChr
```

mean number of window positions with no site differences: `r NoSites$mean_sites`

Breakdown of site distributions

```{r, echo=FALSE}
ZeroVsNonZero
```
That was the end of everything I could get for trouble shooting why there are so many zeros.

Since weighted is needed going to go ahead and do that
```{r Adding Weight, message=FALSE, warning=FALSE}
#weighting it out
weighted_pi <- pi %>%
  group_by(pop, chr_group, subgenome) %>%
  summarize(
    weighted_pi =
      sum(count_diffs, na.rm = TRUE) /
      sum(count_comparisons, na.rm = TRUE),
    .groups = "drop"
  )
```

Pat's paper and I think the cotton paper both used the mean and CI plots. I think the only way to do this with the weighted thing is to make confidence intervals. If there is a better way I would rather do that. I couldn't remember which lab in 830 we did them. But I would happily adjust the code to that from 830 or 831.

```{r BootstrapFx, message=FALSE, warning=FALSE}
boot_pi <- function(df){
  sampled <- df[sample(nrow(df), replace = TRUE), ]
  comps <- sum(sampled$count_comparisons, na.rm = TRUE)
  if(comps == 0) return(NA)
  sum(sampled$count_diffs, na.rm = TRUE) / comps
}
```


```{r FunctionTest, message=FALSE, warning=FALSE, echo=FALSE}
#testing
chr1 <- pi %>%
  filter(chr_group == "chr1")
boot_N <- replicate(
  1000,
  boot_pi(filter(chr1, subgenome == "N"))
)
boot_S <- replicate(
  1000,
  boot_pi(filter(chr1, subgenome == "S"))
)
boot_diff <- boot_N - boot_S
quantile(
  boot_diff,
  c(0.025, 0.975)
)
```

So now that I think the function is working properly. If I even needed to make one in the first place. 

```{r WeightedPi, message=TRUE, warning=TRUE}
#officially making the confidence intrevals
#weighted
weighted_pi_ci <- pi %>%
  group_by(pop, chr_group, subgenome) %>%
  group_modify(~{
    boots <- replicate(
      1000,
      boot_pi(.x)
    )
    tibble(
      weighted_pi =
        sum(.x$count_diffs, na.rm = TRUE) /
        sum(.x$count_comparisons, na.rm = TRUE),
      lower = quantile(boots, 0.025, na.rm = TRUE),
      upper = quantile(boots, 0.975, na.rm = TRUE)
    )
  })
```

``` {r} 
#checking with the plot
#the overall one
ggplot(
  weighted_pi_ci,
  aes(
    chr_group,
    weighted_pi,
    color = pop,
    shape = subgenome
  )
) +
  geom_point(
    position = position_dodge(width = 0.5),
    size = 3
  ) +
  geom_errorbar(
    aes(
      ymin = lower,
      ymax = upper
    ),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  theme_bw() +
  ylab(expression(pi[weighted])) +
  xlab("Chromosome")
#New Colors and Labels
FinalSumPlot <- ggplot(
  weighted_pi_ci,
  aes(
    chr_group,
    weighted_pi,
    color = pop,
    shape = subgenome
  )
) +
  geom_point(
    position = position_dodge(width = 0.5),
    size = 3
  ) +
  geom_errorbar(
    aes(
      ymin = lower,
      ymax = upper
    ),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  scale_color_manual(
    values = pop_colors,
    labels = pop_labels,
    name = "Population"
  ) +
  theme_bw() +
  ylab(expression(pi[weighted])) +
  xlab("Chromosome")

print(FinalSumPlot)

#more of a remake of the first plot
ggplot(weighted_pi_ci, aes(pop,
    weighted_pi,
    color = subgenome)) +
  geom_boxplot() +
  ylab(expression(pi[weighted])) +
  xlab("Population") +
  theme_bw() +
  
#trying something funky
ggplot(weighted_pi_ci, aes(pop,
    weighted_pi,
    color = chr_group,
    shape = subgenome)) +
    geom_point(
    position = position_dodge(width = 0.5),
    size = 3
  ) +
  geom_errorbar(
    aes(
      ymin = lower,
      ymax = upper
    ),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  ylab(expression(pi[weighted])) +
  xlab("Population") +
  theme_bw()
```

Pat's paper used the wilcox test
So going to do pairwise between the subgenomes for chromosomes and populations

```{r Wilcox Subgenome}
wilcox.test(weighted_pi$weighted_pi ~ weighted_pi$subgenome)
wilcox.test(pi$avg_pi ~ pi$subgenome)
#running more than one at a time
#weighted
weighted_subgenome_wilcox <- weighted_pi %>%
  filter(!is.na(weighted_pi)) %>%
  group_by(chr_group) %>%
  group_modify(~{
    test <- wilcox.test(
      weighted_pi ~ subgenome,
      data = .x,
      conf.int = TRUE
    )
    broom::tidy(test)
  })
weighted_subgenome_wilcox
#weighted but sub x population
weighted_subpop_wilcox <- weighted_pi %>%
  filter(!is.na(weighted_pi)) %>%
  group_by(pop) %>%
  group_modify(~{
    test <- wilcox.test(
      weighted_pi ~ subgenome,
      data = .x,
      conf.int = TRUE
    )
    broom::tidy(test)
  })
weighted_subpop_wilcox
#sub x pop x chr doesn't have enough when weighted to run this
```

Getting a better visualization of the subgenome differences. 
```{r}
sub_diff <- pi %>%
  group_by(pop, chr_group, subgenome) %>%
  summarize(
    weighted_pi =
      sum(count_diffs, na.rm = TRUE) /
      sum(count_comparisons, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = subgenome,
    values_from = weighted_pi
  ) %>%
  mutate(
    diff = N - S
  )

sub_diff %>%
  arrange(desc(abs(diff)))

sub_diff_boot <- pi %>%
  group_by(pop, chr_group) %>%
  group_modify(~{
    N_dat <- filter(.x, subgenome == "N")
    S_dat <- filter(.x, subgenome == "S")
    diffs <- replicate(1000, {
      boot_pi(N_dat) - boot_pi(S_dat)
    })
    tibble(
      lower = quantile(diffs, 0.025, na.rm = TRUE),
      upper = quantile(diffs, 0.975, na.rm = TRUE)
    )
  })

sub_diff <- sub_diff %>%
  left_join(sub_diff_boot,
            by = c("pop", "chr_group"))

ggplot(sub_diff,
       aes(chr_group, diff,
           color = pop,
           group = pop)) +
  geom_hline(yintercept = 0,
             linetype = 2) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(ymin = lower,
        ymax = upper),
    width = 0.15
  ) +
  theme_bw() +
  ylab(expression(pi[B.laevigata] - pi[B.stricta])) +
  xlab("Chromosome")

ggplot(sub_diff, aes(pop,
    diff,
    color = chr_group)) +
    geom_point(
    position = position_dodge(width = 0.5),
    size = 3
  ) +
  geom_errorbar(
    aes(
      ymin = lower,
      ymax = upper
    ),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  ylab(expression(pi[B.laevigata] - pi[B.stricta])) +
  xlab("Population") +
  theme_bw()
```
Now time for populations. Since there are five I either need to do something else or do wilcox for each pair. 

```{r Pairwise Pops, message=FALSE, warning=TRUE}
#lets see if this works
#weighted
pairwise.wilcox.test(
  weighted_pi$weighted_pi,
  weighted_pi$pop,
  p.adjust.method = "BH"
)
#not weighted
pairwise.wilcox.test(
  pi$avg_pi,
  pi$pop,
  p.adjust.method = "BH"
)
#no zeros
NoZ <- pi %>%
  filter(avg_pi != "0")
pairwise.wilcox.test(
  NoZ$avg_pi,
  NoZ$pop,
  p.adjust.method = "BH"
)

```

Looking into population relationships with everything else
```{r}
#weighted
PopChr_wilcox <- weighted_pi %>%
  group_by(chr_group) %>%
  group_modify(~{
    test <- pairwise.wilcox.test(weighted_pi$weighted_pi,
                        weighted_pi$pop, 
                        p.adjust.method = "BH"
                        )
    broom::tidy(test)
  })
PopChr_wilcox

PopSub_wilcox <- weighted_pi %>%
  group_by(subgenome) %>%
  group_modify(~{
    test <- pairwise.wilcox.test(weighted_pi$weighted_pi,
                        weighted_pi$pop, 
                        p.adjust.method = "BH"
                        )
    broom::tidy(test)
  })
PopSub_wilcox

PopSubChr_wilcox <- weighted_pi %>%
  group_by(subgenome, chr_group) %>%
  group_modify(~{
    test <- pairwise.wilcox.test(weighted_pi$weighted_pi,
                        weighted_pi$pop, 
                        p.adjust.method = "BH"
                        )
    broom::tidy(test)
  })
PopSubChr_wilcox
```

not weighted T-test from mess up

```{r}
pairwise_pop <- pi %>%
  filter(!is.na(avg_pi)) %>%
  group_by(chr_group) %>%
  group_modify(~{

    pw <- pairwise.t.test(
      .x$avg_pi,
      .x$pop,
      p.adjust.method = "BH"
    )

    as.data.frame(as.table(pw$p.value))

  })

pop_means <- pi %>%
  group_by(pop, chr_group) %>%
  summarize(
    mean_pi = mean(avg_pi, na.rm = TRUE),
    n_windows = sum(!is.na(avg_pi)),
    .groups = "drop"
  )

```

```{r}
pi %>%
  group_by(pop) %>%
  summarize(
    n_windows = n(),
    n_na = sum(is.na(avg_pi)),
    prop_na = mean(is.na(avg_pi))
  )

pi %>%
  group_by(pop, chromosome) %>%
  summarize(
    prop_na = mean(is.na(avg_pi)),
    .groups = "drop"
  )
```

I did like the heat map from before so gonna try that again

```{r}
#weighted
df <- as.data.frame(weighted_pi)

heat_df <- expand_grid(
  pop1 = df$pop,
  pop2 = df$pop
) %>%
  left_join(
    df[, c("pop", "weighted_pi")],
    by = c("pop1" = "pop")
  ) %>%
  rename(pi1 = weighted_pi) %>%
  left_join(
    df[, c("pop", "weighted_pi")],
    by = c("pop2" = "pop")
  ) %>%
  rename(pi2 = weighted_pi) %>%
  mutate(diff = pi1 - pi2)

ggplot(heat_df,
       aes(x = pop1,
           y = pop2,
           fill = diff)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_bw()

```

something is weird witht the heat map assuming I did not do it correctly at this point. 

Population and Chromosomes - not weighted

```{r}
pi_sum <- pi %>%
  group_by(pop, chr_group) %>%
  summarize(
    mean_pi = mean(avg_pi, na.rm = TRUE),
    sd = sd(avg_pi, na.rm = TRUE),
    n = sum(!is.na(avg_pi)),
    se = sd/sqrt(n),
    lower = mean_pi - 1.96 * se,
    upper = mean_pi + 1.96 * se,
    .groups = "drop"
  )

ggplot(
  pi_sum,
  aes(chr_group,
      mean_pi,
      color = pop,
      group = pop)
) +
  geom_point(size = 2) +
  geom_errorbar(
    aes(ymin = lower,
        ymax = upper),
    width = .15
  ) +
  theme_bw()
```

